\vspace{-\baselineskip} Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3.2\baselineskip}

### 2017-03-23 paper/ms.tex
fixing #78 - Equation A4
adding references to Mcquillan 2013, 2014

Another source of variability that can be measured from time series
measurements of stars is rotation.
The inhomogeneous surface of the star (spots, plage, \etc) imprints itself as
quasiperiodic variations in photometric or spectroscopic observations
\citep{Dumusque:2014}.
It has been demonstrated that for light curves with nearly uniform sampling,
the empirical autocorrelation function provides a reliable estimate of the
rotation period of a star \citep{Aigrain:2015} and that a GP model with a
quasiperiodic covariance function can be used to make probabilistic
measurements even with sparsely sampled data (R.~Angus, \etal\ in prep.).
The covariance function used for this type of analysis has the form
\begin{eqnarray}\eqlabel{sine2}
k(\tau) = A\,\exp\left(-\frac{\tau^2}{2\,\ell^2} -
    \Gamma\,\sin^2\left(\frac{\pi\,\tau}{P} \right) \right)
\end{eqnarray}
where $P$ is the period of the oscillation.
The key difference between this function and other quasiperiodic kernels is
that it is positive everywhere.
We can construct a simple \celerite\ covariance function with similar properties
as follows
\begin{eqnarray}\eqlabel{rot-kernel}
k(\tau) = \frac{a}{2+b}\,e^{-c\,\tau}\,\left[
    \cos\left(\frac{2\,\pi\,\tau}{P}\right) + (1 + b)
\right]
\end{eqnarray}
for $a>0$, $b>0$, and $c>0$.
The covariance function in \eq{rot-kernel} cannot exactly reproduce \eq{sine2}
but, since \eq{sine2} is only an effective model, \eq{rot-kernel} can be used
as a drop-in replacement for a substantial gain in computational efficiency.

@@

Using \eq{celerite-psd}, we find that for a single \celerite\ term, this
requirement is met when
\begin{eqnarray}\eqlabel{psd-req}
\frac{(a_j\,c_j+b_j\,d_j)\,({c_j}^2+{d_j}^2)+(a_j\,c_j-b_j\,d_j)\,\omega^2}
{\omega^4+2\,({c_j}^2-{d_j}^2)\,\omega^2+{({c_j}^2+{d_j}^2)}^2} > 0 \quad.
\end{eqnarray}
The denominator is positive for all $c_j \ne 0$ and it can be shown that, when
$c_j=0$, \eq{psd-req} is satisfied for all $\omega \ne d_j$ where the power is
identically zero.
Therefore, when $c_j \ne 0$, we require that the numerator is positive for all
$\omega$.
This requirement can also be written as
\begin{eqnarray}
a_j\,c_j > -b_j\,d_j \\
a_j\,c_j > b_j\,d_j \quad.
\end{eqnarray}
Furthermore, we can see that $a_j$ must be positive since $k(0) = a_j$ should
should be positive and, similarly, by requiring the covariance to be finite at
infinite lag, we obtain the constraint $c_j \ge 0$.
Combining these results, we find the constraint
\begin{eqnarray}
    -|b_j\,d_j| < a_j\,c_j < |b_j\,d_j| \quad.
\end{eqnarray}

>>>>

Another source of variability that can be measured from time series
measurements of stars is rotation.
The inhomogeneous surface of the star (spots, plage, \etc) imprints itself as
quasiperiodic variations in photometric or spectroscopic observations
\citep{Dumusque:2014}.
It has been demonstrated that for light curves with nearly uniform sampling,
the empirical autocorrelation function provides a reliable estimate of the
rotation period of a star \citep{Mcquillan:2013, Mcquillan:2014, Aigrain:2015}
and that a GP model with a
quasiperiodic covariance function can be used to make probabilistic
measurements even with sparsely sampled data (R.~Angus, \etal\ in prep.).
The covariance function used for this type of analysis has the form
\begin{eqnarray}\eqlabel{sine2}
k(\tau) = A\,\exp\left(-\frac{\tau^2}{2\,\ell^2} -
    \Gamma\,\sin^2\left(\frac{\pi\,\tau}{P} \right) \right)
\end{eqnarray}
where $P$ is the period of the oscillation.
The key difference between this function and other quasiperiodic kernels is
that it is positive everywhere.
We can construct a simple \celerite\ covariance function with similar properties
as follows
\begin{eqnarray}\eqlabel{rot-kernel}
k(\tau) = \frac{a}{2+b}\,e^{-c\,\tau}\,\left[
    \cos\left(\frac{2\,\pi\,\tau}{P}\right) + (1 + b)
\right]
\end{eqnarray}
for $a>0$, $b>0$, and $c>0$.
The covariance function in \eq{rot-kernel} cannot exactly reproduce \eq{sine2}
but, since \eq{sine2} is only an effective model, \eq{rot-kernel} can be used
as a drop-in replacement for a substantial gain in computational efficiency.

@@

Using \eq{celerite-psd}, we find that for a single \celerite\ term, this
requirement is met when
\begin{eqnarray}\eqlabel{psd-req}
\frac{(a_j\,c_j+b_j\,d_j)\,({c_j}^2+{d_j}^2)+(a_j\,c_j-b_j\,d_j)\,\omega^2}
{\omega^4+2\,({c_j}^2-{d_j}^2)\,\omega^2+{({c_j}^2+{d_j}^2)}^2} > 0 \quad.
\end{eqnarray}
The denominator is positive for all $c_j \ne 0$ and it can be shown that, when
$c_j=0$, \eq{psd-req} is satisfied for all $\omega \ne d_j$ where the power is
identically zero.
Therefore, when $c_j \ne 0$, we require that the numerator is positive for all
$\omega$.
This requirement can also be written as
\begin{eqnarray}
a_j\,c_j > -b_j\,d_j \\
a_j\,c_j > b_j\,d_j \quad.
\end{eqnarray}
Furthermore, we can see that $a_j$ must be positive since $k(0) = a_j$ should
should be positive and, similarly, by requiring the covariance to be finite at
infinite lag, we obtain the constraint $c_j \ge 0$.
Combining these results, we find the constraint
\begin{eqnarray}
    |b_j\,d_j| < a_j\,c_j \quad.
\end{eqnarray}
